
In this file, we're going to load the tidy data created in tidy_coral_data.Rmd and extract information from the note columns to populate new dynamic columns.

The dynamic columns we'll make are death, recruitment, growth, fission, fusion, and f-group.

## Setup
```{r}
# Clean workspace
rm(list = ls())

# Load necessary libraries
library(tidyverse)
library(here)
library(janitor)

# Load our tidy coral dataset
coral_df_tidy <- read_csv(here("data", "coral_tidy_2024.csv"))
```
## Aggregating and tokenizing notes

```{r}
# Let's extract the note values for each observation and place them in a list in each column
# we'll have a temporary field called notes_extracted that will hold the individual tokenized notes as a list

# To extract the notes, take the column note and note_extra, splice each along commas and semicolons, trim, and convert to lowercase. Combine the lists from note and note_extra to get the value for notes_extracted


# coral_df_tidy <- coral_df_tidy %>% 
#   mutate(notes_extracted = map2(note, note_extra, ~c(strsplit(tolower(.x), "[,;]")[[1]], strsplit(tolower(.y), "[,;]")[[1]])))

# Rewrite, but also use str_trim
coral_df_tidy <- coral_df_tidy %>% 
  mutate(notes_extracted = map2(note, note_extra, ~c(strsplit(tolower(.x), "[,;]")[[1]], strsplit(tolower(.y), "[,;]")[[1]]))) %>% 
  mutate(notes_extracted = map(notes_extracted, ~str_trim(.x)))


```

## Parsing notes for dynamics

### Death
```{r}
# Initialize empty death column
coral_df_tidy <- coral_df_tidy %>% mutate(death = 0)

# If notes_extracted contains a "d", then set the new column "death" to 1. 
# It should be 0 otherwise.
coral_df_tidy <- coral_df_tidy %>% 
  rowwise() %>%
  mutate(death = ifelse("d" %in% unlist(notes_extracted), 1, 0)) %>%
  ungroup()
```

### Growth
```{r}
# Initialize empty growth column
coral_df_tidy <- coral_df_tidy %>% mutate(growth = 0)

# Same as above, but for the value g
coral_df_tidy <- coral_df_tidy %>% 
  rowwise() %>%
  mutate(growth = ifelse("g" %in% unlist(notes_extracted), 1, 0)) %>%
  ungroup()
```


### Recruitment
```{r}
# First initialize an empty recruitment column
coral_df_tidy <- coral_df_tidy %>% mutate(recruitment = 0)

# If notes_extracted contains "r", "(r)", "r-fr" then set the  new column 
# recruitment to 1. It should be 0 otherwise
for (i in 1:nrow(coral_df_tidy)) {
  note <- coral_df_tidy$notes_extracted[[i]]
  if ("r" %in% note || "(r)" %in% note || "r-fr" %in% note) {
    coral_df_tidy$recruitment[i] <- 1
  }
}

```

### Fusion
We know that fusion ocurred at an observation
- if the token itself EQUALS 'fuX', 'fusX', 'fuX-ed' where X consists of zero or more digits
- if the token itself contains the word "fused" or "fusion"

In the first case, the string of digits X should be saved in a field called fusion_group


```{r}
# Start by initializing empty columns fusion and fusion_group
coral_df_tidy <- coral_df_tidy %>% 
  mutate(fusion = 0, fusion_group = NA)

for(i in 1:nrow(coral_df_tidy)) {
  notes <- coral_df_tidy$notes_extracted[[i]]
  
  # Iterate through all values in notes
  for (note in notes) {
    # print(paste("Note is ", note))
    if (grepl("^fu[0-9]*$", note) || 
        grepl("^fus[0-9]*$", note) || 
        grepl("^fu[0-9]*-ed$", note)) {
      # This is the ideal case where we have a fuX or fusX or fuX-ed.
      # We'll be able to extract a fusion_group number from this format
      coral_df_tidy$fusion[i] <- 1
      fus_group <- str_extract(note, "^fus?([0-9]*)(-ed)?$", group = 1)
      coral_df_tidy$fusion_group[i] <- as.numeric(fus_group)
    } else if(grepl("fused", note) || grepl("fusion", note)) {
      # It's possible that we'll get these sorts of notes, but we shouldn't
      # overwrite any existing groups, that's why if fusion has already been set
      # we'll just leave it as is.
      if(coral_df_tidy$fusion[i] == 1) {
        next
      }
      coral_df_tidy$fusion[i] <- 1
      coral_df_tidy$fusion_group[i] <- NA
    }
  }
}

```


### Fission
This is similar to the fusion case, but with less possibilities
We label fission as having ocurred if
- if the token itself EQUALS 'fiX' or 'fisX' where consists of zero or more digits
- if the token contains the word 'fission'

```{r}
# Start by initializing empty columns fission and fission_group
coral_df_tidy <- coral_df_tidy %>% 
  mutate(fission = 0, fission_group = NA)

for(i in 1:nrow(coral_df_tidy)) {
  notes <- coral_df_tidy$notes_extracted[[i]]
  
  # Iterate through all values in notes
  for (note in notes) {
    if (grepl("^fis?[0-9]*$", note)) {
      # This is the ideal case where we have a fiX or fisX.
      # We'll be able to extract a fission_group number from this format
      coral_df_tidy$fission[i] <- 1
      fis_group <- str_extract(note, "^fis?([0-9]*)$", group = 1)
      coral_df_tidy$fission_group[i] <- as.numeric(fis_group)
    } else if(grepl("fission", note)) {
      # It's possible that we'll get these sorts of notes, but we shouldn't
      # overwrite any existing groups, that's why if fission has already been set
      # we'll just leave it as is.
      if(coral_df_tidy$fission[i] == 1) {
        next
      }
      
      
      
      coral_df_tidy$fission[i] <- 1
      coral_df_tidy$fission_group[i] <- NA
    }
  }
}


```


## Storage
```{r}
# remove the notes_extracted column, then save it as a csv
# let's call it "coral_tidy_2024_with_dynamics.csv"
coral_df_tidy <- coral_df_tidy %>% select(-notes_extracted) %>% 
  write_csv(here("data", "coral_tidy_2024_with_dynamics.csv"))
```




