

```{r}
# Clean env
rm(list = ls())

# Load necessary libraries
library(tidyverse)
library(here)
library(janitor)
library(readxl)

# Load coral data from excel file
# coral_df_raw <- read_excel(here("data", "coral_raw_wide_2024.xlsx")) %>% 
#   clean_names()
# ^ Commented out for the time-being, all our aggregated data is spread out through other sheets
# We'll need to load all the data from these individual sheets and combine them with each other
# in order to obtain a coral_df_raw

# Get the column names of each sheet in the file that matches the pattern "LTERXAA"
sheets_habitats <- excel_sheets(here("data", "coral_raw_wide_2024.xlsx")) %>% 
  str_subset("^LTER[0-9]{1}[A-Z]{2}$")

COLUMN_LIMIT <- 100 # The maximum number of columns we'll need to load from each sheet.
                    # May need to be tweaked as dataset grows

# Load each of the sheets and join them togethr
# by stacking the datasets on top of each other.
# The columns have significant overlap, but there might be some unique ones
# among the individual sheets
# Convert each sheet to be all string type to avoid type conflicts
coral_df_raw <- sheets_habitats %>%
  map(~read_excel(here("data", "coral_raw_wide_2024.xlsx"), sheet = .x, range = cell_cols(c(1,COLUMN_LIMIT)))) %>%
  map(~mutate_all(., as.character)) %>%
  map(~clean_names(.)) %>%
  map( # Remove underscores from column names
    ~rename_with(., ~str_remove(., "_"), everything())
  ) %>%
  reduce(bind_rows)

# coral_df_raw <- sheets_habitats %>% 
#   map(~read_excel(here("data", "coral_raw_wide_2024.xlsx"), sheet = .x, range = cell_cols(c(1,90)))) %>%
#   map(~mutate_all(., as.character)) %>%
#   map(~clean_names(.)) %>% 
#   map( # Remove underscores from column names
#     ~rename_with(., ~str_remove(., "_"), everything())
#   )

# Get col names of each element in coral_df_raw
# coral_df_raw %>% map(~colnames(.))

# do again, but get the name of each sheet alongside the columns
# coral_df_raw %>% 
#   map(~colnames(.)) %>% 
#   map2(sheets_habitats, ~set_names(.x, .y))

# sheets_habitats %>% 
#   map(~read_excel(here("data", "coral_raw_wide_2024.xlsx"), sheet = .x) %>% colnames())
```

## Tidying Data

There are a lot of columns we don't need
```{r}
# First, we only want to select the columns we're interested in:
# site, habitat, transect, taxa, x, y, z, lYYYY, wYYY, hYYYY, obsYYYY, note_YYYY, note_YYYY_extra
# Make sure that for the column names with years, the exact format is followed.
# We'll also add a coral_number field for each row of the raw data frame, just so we can track each coral by its distinct identity
coral_df <- coral_df_raw %>% 
  mutate(coral_number = row_number()) %>%
  select(coral_number, site, habitat, transect, taxa, x, y, z, 
         matches("l[0-9]{4}"), 
         matches("w[0-9]{4}"), 
         matches("h[0-9]{4}"), 
         matches("obs[0-9]{4}"), 
         matches("note_[0-9]{4}"), 
         matches("note_[0-9]{4}_extra"))

# Because it'll make things a little easier down the line, let's get the note_YYYY_extra columns to instead use the format note_extra_YYYY (ensuring that the col name ends in a year)
coral_df <- coral_df %>% 
  rename_with(~str_replace(., "note_([0-9]{4})_extra", "note_extra_\\1"), matches("note_[0-9]{4}_extra"))

```




There are two note columns, note_YYYY and note_YYYY_extra. Later we'll deal with aggregating these two.

Let's tidy this data. We'll want to have a year column. We'll also want to have a column for each of length, width, height, observer, and note. 

```{r}

# First convert anything numeric to be a string to avoid type conflicts
coral_df <- coral_df %>% 
  mutate(across(where(is.numeric), as.character))

coral_df_tidy <- coral_df %>%
  pivot_longer(cols = matches("^l\\d{4}$|^w\\d{4}$|^h\\d{4}$|^obs\\d{4}$|^note_\\d{4}$|^note_extra_\\d{4}$"), 
               names_to = "key", 
               values_to = "value") %>%
  mutate(year = str_extract(key, "\\d{4}"))

# Trim off the year from every thing in the key column (including the underscore that sometimes precedes dates)
coral_df_tidy <- coral_df_tidy %>% 
  mutate(key = str_remove(key, "_*\\d{4}"))

# Now pivot wider to get the columns we want
coral_df_tidy <- coral_df_tidy %>%
  pivot_wider(names_from = key, values_from = value) %>%
  rename(length = l, width = w, height = h, observer = obs)
```

So we aren't getting any list-cols! That's good.

Now the last thing we need to do is convert the fields x, y, z, year, length, width, height to be properly numeric again.

```{r}
coral_df_tidy <- coral_df_tidy %>%
  mutate_at(vars(x, y, z, year, length, width, height), as.numeric)
```


Let's store this data file so that we can use it later.

```{r}
write_csv(coral_df_tidy, here("data", "coral_tidy_2024.csv"))
```



