
## Setup
```{r}
# Clear env
rm(list=ls())

# Load necessary libraries
library(tidyverse)
library(here)

# Load data
coral_df <- read_csv(here("data", "coral_tidy_2024_with_dynamics.csv"))
```

## Create aggregated 'dynamic' field

The fields 
"death"         "growth"        "recruitment"  
 "fusion"        "fusion_group"  "fission"       "fission_group"
 are to be aggregated into a single dynamics field.
 
```{r}
# Make empty dynamics field
coral_df <- coral_df %>% mutate(dynamics = NA)

# Loop through each item in coral_df
for(i in 1:nrow(coral_df)){
  # Initialize empty dynamics vector
  dyn <- c()
  
  if(coral_df$death[[i]] == 1) {
    dyn <- append(dyn, "Death")
  }
  if(coral_df$recruitment[[i]] == 1) {
    dyn <- append(dyn, "Recruitment")
  }
  if(coral_df$growth[[i]] == 1) {
    dyn <- append(dyn, "Growth")
  }
  if(coral_df$fission[[i]] == 1) {
    fis_group <- ifelse(is.na(coral_df$fission_group[[i]]), "", coral_df$fission_group[[i]])
    dyn <- append(dyn, paste0("Fission", fis_group))
  }
  if(coral_df$fusion[[i]] == 1) {
    fus_group <- ifelse(is.na(coral_df$fusion_group[[i]]), "", coral_df$fusion_group[[i]])
    dyn <- append(dyn, paste0("Fusion", fus_group))
  }
  
  # Collapse 
  dyn_str <- paste(dyn, collapse = ", ")
  if(dyn_str != "") # Only write new data if string is non-empty (leave as NA otherwise)
    coral_df$dynamics[[i]] <- dyn_str

}
  
```


## Widen the data

First prepare the 'longer' format.
This will consist of key-value-ifying the important columns that will be depend on each year, and appending the year to each of their names.
```{r}
# Remove the dummy variables
coral_df_longer <- coral_df %>% select(-death, -growth, -recruitment, -fusion, -fusion_group, -fission, -fission_group)

year_index <- which(names(coral_df_longer) == "year")
coral_df_longer <- coral_df_longer %>%  mutate(across((year_index+1):last_col(), as.character))

# pivot longer all columns from "year" onwards
#  names go to a column called "key" and the values go to a column called "value"
coral_df_longer <- coral_df_longer %>% pivot_longer(cols = (year_index+1):last_col(), names_to = "key", values_to = "value")

# For append the year to the end of each key
coral_df_longer <- coral_df_longer %>% mutate(key = paste0(key, "_", year))

```

Now we widen the data. We can take the 'longer' df and perform a simple pivot_wider()

```{r}
coral_df_wider <- coral_df_longer %>%
  select(-year) %>% 
  pivot_wider(names_from = key, values_from = value)
```


## Save the data
```{r}
write_csv(coral_df_wider, here("data", "coral_wide_2024_with_standardized_dynamics.csv"))
```