
## Setup
```{r}
# Clear env
rm(list=ls())

# Load necessary libraries
library(tidyverse)
library(here)
library(openxlsx)

# Load data
coral_df <- read_csv(here("data", "coral_tidy_2024_with_dynamics.csv"))
```

## Create aggregated 'dynamic' field

The fields 
 
```{r}
# Make empty dynamics field
coral_df <- coral_df %>% mutate(dynamics = NA)

# Each dynamic field has a name following the convention 'dyn_DYNAMICNAME'
# 
# Loop through each item in coral_df
for(i in 1:nrow(coral_df)){
  # Initialize empty dynamics vector
  dyn <- c()

  if(coral_df$dyn_death[[i]] == 1) {
    dyn <- append(dyn, "Death")
  }
  if(coral_df$dyn_recruitment[[i]] == 1) {
    dyn <- append(dyn, "Recruitment")
  }
  if(coral_df$dyn_growth[[i]] == 1) {
    dyn <- append(dyn, "Growth")
  }
  if(coral_df$dyn_shrinkage[[i]] == 1) {
    dyn <- append(dyn, "Shrinkage")
  }
  if(coral_df$dyn_immigration[[i]] == 1) {
    dyn <- append(dyn, "Immigration")
  }
  if(coral_df$dyn_emigration[[i]] == 1) {
    dyn <- append(dyn, "Emigration")
  }
  if(coral_df$dyn_edge[[i]] == 1) {
    dyn <- append(dyn, "Edge")
  }
  if(coral_df$dyn_fission[[i]] == 1) {
    fis_group <- ifelse(is.na(coral_df$fission_group[[i]]), "", coral_df$fission_group[[i]])
    dyn <- append(dyn, paste0("Fission", fis_group))
  }
  if(coral_df$dyn_fusion[[i]] == 1) {
    fus_group <- ifelse(is.na(coral_df$fusion_group[[i]]), "", coral_df$fusion_group[[i]])
    dyn <- append(dyn, paste0("Fusion", fus_group))
  }


  # Collapse
  dyn_str <- paste(dyn, collapse = ", ")
  if(dyn_str != "") # Only write new data if string is non-empty (leave as NA otherwise)
    coral_df$dynamics[[i]] <- dyn_str

}
  
```


## Widen the data

First prepare the 'longer' format.
This will consist of key-value-ifying the important columns that will be depend on each year, and appending the year to each of their names.
```{r}
# Remove the dummy variables (anything that starts with dyn_)
coral_df_longer <- coral_df %>% 
  select(-starts_with("dyn_")) %>% 
  select(-fusion_group, -fission_group)

year_index <- which(names(coral_df_longer) == "year")
coral_df_longer <- coral_df_longer %>%  mutate(across((year_index+1):last_col(), as.character))

# pivot longer all columns from "year" onwards
#  names go to a column called "key" and the values go to a column called "value"
coral_df_longer <- coral_df_longer %>% pivot_longer(cols = (year_index+1):last_col(), names_to = "key", values_to = "value")

# For append the year to the end of each key
coral_df_longer <- coral_df_longer %>% mutate(key = paste0(key, "_", year))

```

Now we widen the data. We can take the 'longer' df and perform a simple pivot_wider()

```{r}
coral_df_wider <- coral_df_longer %>%
  select(-year) %>% 
  pivot_wider(names_from = key, values_from = value)
```

Since we're done pivoting, we can convert our numeric data back to being properly a numeric type.

We would convert
- length_YYYY
- width_YYYY
- height_YYYY
- growth_ratio_YYYY

```{r}
coral_df_wider <- coral_df_wider %>% 
  mutate_at(vars(matches("length_[0-9]{4}|width_[0-9]{4}|height_[0-9]{4}|growth_ratio_[0-9]{4}")), as.numeric)
```



## Save the data
```{r}
write_csv(coral_df_wider, here("data", "cleaning_outputs", "coral_clean_wide_2024.csv"))
```

## Saving in special format
NOTE this is now done in a dedicated excel script.
Let's save this as a xlsx file with multiple tabs
```{r}
# wb <- createWorkbook()
# addWorksheet(wb, "ALL_SITES")
# writeData(wb, sheet = "ALL_SITES", x = coral_df_wider)
# 
# # Loop through the unique values for site and habitat
# site_habitat <- coral_df_wider %>% select(site, habitat) %>% distinct()
# for(i in 1:nrow(site_habitat)){
#   site <- site_habitat$site[[i]]
#   habitat <- site_habitat$habitat[[i]]
#   sheet_name <- paste0(site, "_", habitat)
#   addWorksheet(wb, sheet_name)
#   writeData(wb, sheet = sheet_name, x = coral_df_wider %>% filter(site == site_habitat$site[[i]], habitat == site_habitat$habitat[[i]]))
# }
# 
# saveWorkbook(wb, file = "coral_demography_2013-2023_cleaned.xlsx", overwrite = TRUE)

```

