
Let's examine Hunter's tweaks to the 2023 data as it compares to the human
friendly one that I generate from our existing data.

data_outputs/coral_clean_wide_human-friendly.xlsx
data_raw/coral_raw_wide_2024_hunter-mod_sept-oct.xlsx

```{r}
# Clean env
rm(list=ls())

# Load necessary libraries
library(tidyverse)
library(here)
library(janitor)
library(readxl)

# My existing datasheet
coral_df <- read_excel(here("data_outputs", "coral_clean_wide_human-friendly.xlsx")) %>% 
  clean_names()

# Hunter's modified data
coral_df_hunter_mod <- read_excel(here("data_raw", "coral_raw_wide_2024_hunter-mod_sept-oct.xlsx")) %>% 
  clean_names()


```
```{r}
# Let's start getting diffs for columns in each. 

# What is the difference between the l2023 field for either, for example

# For example
# Find the indices where the vectors differ
# diff_indices <- which(vec1 != vec2)
# 
# # Display the indices and the corresponding differing values
# data.frame(
#   Index = diff_indices,
#   Vec1_Value = vec1[diff_indices],
#   Vec2_Value = vec2[diff_indices]
# )

# Write a function that takes in 2 equal-length vectors and returns a dataframe
# that contains their diff in this manner
get_diff <- function(vec1, vec2){
  diff_indices <- which(vec1 != vec2)
  data.frame(
    Index = diff_indices,
    Vec1_Value = vec1[diff_indices],
    Vec2_Value = vec2[diff_indices]
  )
}



```

```{r}
# Get the differences for the l2023 field
get_diff(coral_df$l2023, coral_df_hunter_mod$l2023)
```

```{r}
# Get the differences for the w2023 field
get_diff(coral_df$w2023, coral_df_hunter_mod$w2023)
```

```{r}
# Get the differences for the h2023 field
get_diff(coral_df$h2023, coral_df_hunter_mod$h2023)
```

```{r}
# Get the differences for the note_2023 field
get_diff(coral_df$note_2023, coral_df_hunter_mod$note_2023)
```

```{r}
# Get the differences for the note_2023_extra field
get_diff(coral_df$note_2023_extra, coral_df_hunter_mod$note_2023_extra)
```

```{r}
# Get the differences for the dynamics_2023 field
get_diff(coral_df$dynamics_2023, coral_df_hunter_mod$dynamics_2023)
```

Exploration complete. It appears hunter only set Missing Data to Death a bunch of times, but didn't do anything else besides that.

It is not worth keeping this modified file. It does not provide anything else useful. I will add it to the exploration script section for future reference, but remove it from the raw directory.