
In this exploration, we'll produce the following
- [ ]  Plot of recruitment each year for each genera at each site
- [ ]  regressions analyses of growth as a function of size for each genera at each site, grouping 2013-2019 (before bleaching) vs. 2020-24 (after bleaching)
- [ ]  Mortality as a function of size …same as #2.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r environment}
# Clean en√
rm(list=ls())

# Load libraries
library(tidyverse)
library(here)
```


```{r load data}
# Loading data
# Let's get it from data_outputs/coral_tidy_2024_with_dynamics.csv

coral_df <- read_csv(here("data_outputs", "coral_tidy_2024_with_dynamics.csv"))

# Note: there was one coral erroneously entered in with taxa of "A" instead of "Acr", let's quickly fix that
coral_df <- coral_df %>% 
  mutate(taxa = ifelse(taxa == "A", "Acr", taxa))


```

## Color Schemes
A note on color schemes, there is some particularity about which colors correspond to which coral taxa


Poc is pure red (255,0, 0) 
Por 0,168,48
Acr is 11,31,143

Millipora was not present before, so we'll come up with a new color

```{r}
# Define custom colors for coral taxa
coral_colors <- c(
  "Poc" = rgb(255, 0, 0, maxColorValue = 255),      # Pure Red
  "Por" = rgb(0, 168, 48, maxColorValue = 255),     # Green
  "Acr" = rgb(11, 31, 143, maxColorValue = 255),     # Blue
  "Mil" = rgb(255, 165, 0, maxColorValue = 255)    # Orange for Millipora
)
```



Let's now plot recruitment each year for each genera at each site
```{r plot recruitment}

recruitment_df <- coral_df %>% 
  group_by(site, taxa, year) %>% 
  summarize(recruitment = sum(dyn_recruitment))

# Let's now plot this where 
# the color is set to taxa
# the facet is set to site
# recruitment_df %>% 
#   ggplot(aes(x = factor(year), y = recruitment, color = taxa)) +
#   geom_line(aes(group = taxa)) +  # ensure lines connect within taxa
#   geom_point() +
#   facet_wrap(~site) +
#   labs(
#     title = "Recruitment by Year, Taxa, and Site",
#     x = "Year",
#     y = "Number of Recruits"
#   ) +
#   theme_minimal()

# Recreate plot, but with year labels every 2 years instead of every year
recruitment_df %>% 
  ggplot(aes(x = factor(year), y = recruitment, color = taxa)) +
  geom_line(aes(group = taxa)) +  # ensure lines connect within taxa
  geom_point() +
  facet_wrap(~site) +
  scale_x_discrete(breaks = seq(2013, 2024, by = 2)) +  # Set x-axis breaks every 4 years
  labs(
    title = "Recruitment by Year, Taxa, and Site",
    x = "Year",
    y = "Number of Recruits"
  ) +
  theme_minimal()

# Regenarate with coral color scheme
recruitment_df %>% 
  ggplot(aes(x = factor(year), y = recruitment, color = taxa)) +
  geom_line(aes(group = taxa)) +  # ensure lines connect within taxa
  geom_point() +
  facet_wrap(~site) +
  scale_x_discrete(breaks = seq(2013, 2024, by = 2)) +  # Set x-axis breaks every 4 years
  scale_color_manual(values = coral_colors) +  # Use custom coral colors
  labs(
    title = "Recruitment by Year, Taxa, and Site",
    x = "Year",
    y = "Number of Recruits"
  ) +
  theme_minimal()

```

## Growth as a function of coral size (ellipsoid model)

To match the pattern in the previous paper, we want to be able to quantify the surface area of each coral
So let's go ahead then and do that.

Let's add a calculated field for surface area , treating the coral colony as an ellipsoid

```{r}
# Create calculated field, call it ellipsoid_surface_area
# derive from length, width, height
coral_df <- coral_df %>%
  mutate(
    length = as.numeric(length),
    width = as.numeric(width),
    height = as.numeric(height),
    ellipsoid_surface_area = 4 * pi * (( (length * width)^1.6 + (length * height)^1.6 + (width * height)^1.6 ) / 3)^(1/1.6)
  )

```

We've also thus far been using growth_ratio as a quotient of averages across dimensions, so we can also make a size quantifier with avg dimension

```{r}
coral_df <- coral_df %>%
  mutate(
    avg_dimension = (length + width + height) / 3
  )
```

```{r}
# Let's go ahead now and make that size plot
# Let's facet by site, color by taxa, include gaps of 2 between years
coral_df %>%
  filter(!is.na(growth_ratio) & growth_ratio > 0 & growth_ratio < 80 & !is.na(avg_dimension) & avg_dimension > 0) %>%
  ggplot(aes(x = avg_dimension, y = growth_ratio, color = taxa)) +
  geom_point(alpha = 0.6) +
  # geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line
  facet_wrap(~site) +
  scale_color_manual(values = coral_colors) +  # Use custom coral colors
  labs(
    title = "Growth Ratio vs. Average Dimension by Site and Taxa",
    x = "Average Dimension (cm)",
    y = "Growth Ratio"
  ) +
  theme_minimal()

```

Dang that's not telling us a whole lot.

Let's try using the ellipsoid surface area on the x

```{r}
coral_df %>%
  filter(!is.na(growth_ratio) & growth_ratio > 0 & growth_ratio < 80 & !is.na(ellipsoid_surface_area) & ellipsoid_surface_area > 0) %>%
  ggplot(aes(x = ellipsoid_surface_area, y = growth_ratio, color = taxa)) +
  geom_point(alpha = 0.6) +
  # geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line
  facet_wrap(~site) +
  scale_color_manual(values = coral_colors) +  # Use custom coral colors
  labs(
    title = "Growth Ratio vs. Ellipsoid Surface Area by Site and Taxa",
    x = "Ellipsoid Surface Area (cm²)",
    y = "Growth Ratio"
  ) +
  theme_minimal()
```



Eh not telling us much still. Let's leave this one alone for now.


## Mortality

So first off lets graph dyn_death as a function of size, across the sites. Color by species.
dyn_death will be 0 or 1, so we need some vertical and horizontal jitter to make them more visible.

We shouldn't be including every single observation here, only concrete instances when a coral was observed. or when an absence of it was observed, and that absence is labeled as death

ALERT: rows that have dyn_death frequently lack length, width, height values. Thus the data must be transformed to include the previously measured dimensions in observations that are indicative of mortality.

```{r}
coral_df %>%
  filter(dyn_death == 1 | (dyn_death == 0 & !is.na(length) & !is.na(width) & !is.na(height))) %>%
  ggplot(aes(x = ellipsoid_surface_area, y = dyn_death, color = taxa)) +
  geom_jitter(width = 0, height = 0.1, alpha = 0.6, size=0.001) +
  facet_wrap(~site) +
  scale_color_manual(values = coral_colors) +  # Use custom coral colors
  scale_x_log10(labels = scales::scientific) +  # Logarithmic x-axis with scientific notation
  labs(
    title = "Mortality (dyn_death) vs. Surface Area by Site and Taxa",
    x = "Surface Area (cm², log scale)",
    y = "Mortality (dyn_death)"
  ) +
  theme_minimal()
```







