
In this notebook, I intend to create an LME model for predicting coral size
in year t+1 based on coral size in year t, while accounting for random effects
based on the transect sampled.

```{r setup}
# Clean Env
rm(list=ls())
# Load libs
library(tidyverse)
library(here)
# load lme library
library(nlme)

# Load data
coral_df <- read_csv(here("data_outputs", "coral_tidy_2024_with_dynamics.csv"))
```

First off we're going to need to create a size calculated field.
We quantify size as the average of the 3 dimensions.

```{r}
# real quick convert the length, width, height columns to numeric
coral_df <- coral_df %>%
  mutate(across(c(length, width, height), as.numeric))

coral_df <- coral_df %>%
  mutate(size = (length + width + height) / 3)

```

Next we're going to need to have each coral observation not only have its size
in the current observation, but also its previously observed size (if one does exist)

To do this, we're going to need to group by transect and coral_number (it's unique identifier), then arrange by year,
then use the lag function to get the previous size.

```{r}
coral_df_lagged <- coral_df %>%
  group_by(transect, coral_number) %>%
  arrange(year) %>%
  mutate(prev_size = lag(size)) %>%
  ungroup()
```

Let's verify that this worked

```{r}
# Look at a specific transect and coral number
# Let's see in year 2014, coral_number=LTER1_BR_P01_Acr_4.15_50_53,
# this coral shrank from 2013
coral_df_lagged %>%
  filter(coral_number == "LTER1_BR_P01_Acr_4.15_50_53") %>%
  select(year, size, prev_size)

# Looks good!

```

One more step we'll need to take before creating the lme model:
- Some transects have the same names across sites, therefore we
need to pair the transect with the site to create a unique identifier.
Let's call this site_transect

```{r}
coral_df_lagged <- coral_df_lagged %>%
  mutate(site_transect = str_c(site, transect, sep = "_"))
```

Now we can go ahead with creating our lme model.
The random-effect will be site_transect

We'll also want to separate our data into 2 groups to run the model on:
1) 2013-2019 (prior to mass bleaching event)
2) 2020 and onwards (after the bleaching event)

```{r}
# Filter out to only include rows where both size and prev_size are not NA
coral_df_for_lme <- coral_df_lagged %>%
  filter(!is.na(prev_size) & !is.na(size))

coral_lme_pre_bleaching <- coral_df_for_lme %>%
  filter(year <= 2019)

coral_lme_post_bleaching <- coral_df_for_lme %>%
  filter(year >= 2020)


# Create the lme model
lme_model_pre <- lme(size ~ prev_size, random = ~1 | site_transect, data = coral_lme_pre_bleaching)
lme_model_post <- lme(size ~ prev_size, random = ~1 | site_transect, data = coral_lme_post_bleaching)
```

Now let's create a visualization to compare the data and model results
for the pre and post bleaching years

```{r}

coral_colors <- c(
  "Poc" = rgb(255, 0, 0, maxColorValue = 255),      # Pure Red
  "Por" = rgb(0, 168, 48, maxColorValue = 255),     # Green
  "Acr" = rgb(11, 31, 143, maxColorValue = 255),     # Blue
  "Mil" = rgb(255, 165, 0, maxColorValue = 255)    # Orange for Millipora
)

# Create a function to plot the data and model results
plot_lme_results <- function(data, model, title) {
  # Create a sequence of prev_size values for predictions
  prev_size_seq <- seq(min(data$prev_size, na.rm = TRUE), max(data$prev_size, na.rm = TRUE), length.out = 100)
  
  # Create a new data frame for predictions
  pred_data <- data.frame(prev_size = prev_size_seq)
  
  # Get the fixed effects predictions
  pred_data$predicted_size <- predict(model, newdata = pred_data, level = 0) # Here level=0 means only fixed effects are used in model outputs.
  
  # Plot the data and the model predictions
  ggplot(data, aes(x = prev_size, y = size, color = taxa)) +
    geom_point(alpha = 0.5) +
    geom_line(data = pred_data, aes(x = prev_size, y = predicted_size), color = "blue", size = 1) +
    scale_x_log10(labels = scales::scientific) +
    scale_y_log10(labels = scales::scientific) +
    scale_color_manual(values = coral_colors) +
    labs(title = title, x = "Size at Year y", y = "Size at Year y+1") +
    theme_minimal()
  
   
}

# Plot pre-bleaching results
plot_lme_results(coral_lme_pre_bleaching, lme_model_pre, "Pre-Bleaching (2013-2019)")

# Plot post-bleaching results
plot_lme_results(coral_lme_post_bleaching, lme_model_post, "Post-Bleaching (2020-2024)")
```
