**New Needs**

- Show Recruitment as a per unit area (each transect is 5 square meters)
- Separate graphs for Habitat (Lagoon vs Outer Slope) — or BR vs OR
- Mohsen: plot the 4 replicate recruitment densities for each area (1 for each transect), and then have a curve representing the mean + a confidence interval.
- You would do this for site and habitat.
- (Possible Concern) Ensuring that individuals fissioning into smaller ones (like Por) gets shown…. though maybe this ought not to be considered recruitment
- Mohsen: says to just take into account the 4 transects T01...T04 for each. ignore the "plots" (P01, P02, etc)


```{r environment}
# Clean en√
rm(list=ls())

# Load libraries
library(tidyverse)
library(here)
```


```{r load data}
# Loading data
# Let's get it from data_outputs/coral_tidy_2024_with_dynamics.csv

coral_df <- read_csv(here("data_outputs", "coral_tidy_2024_with_dynamics.csv"))

# Note: there was one coral erroneously entered in with taxa of "A" instead of "Acr", let's quickly fix that
coral_df <- coral_df %>% 
  mutate(taxa = ifelse(taxa == "A", "Acr", taxa))


```



## Color Schemes
A note on color schemes, there is some particularity about which colors correspond to which coral taxa


Poc is pure red (255,0, 0) 
Por 0,168,48
Acr is 11,31,143
Millipora is orange

```{r color-scheme}
# Define custom colors for coral taxa
coral_colors <- c(
  "Poc" = rgb(255, 0, 0, maxColorValue = 255),      # Pure Red
  "Por" = rgb(0, 168, 48, maxColorValue = 255),     # Green
  "Acr" = rgb(11, 31, 143, maxColorValue = 255),     # Blue
  "Mil" = rgb(255, 165, 0, maxColorValue = 255)    # Orange for Millipora
)
```

## Recruitment Summarization

- get the recruitment grouped by site, transect, taxa, habitat, and year
- Then we'll sum dyn_recruitment for each group to get a raw count of recruits
- Then we'll filter the resulting data frame to keep only transects that are named T01, T02, T03, or T04
- Then we'll create a new column called recruitment_density which is recruitment divided by 5 (since each transect is 5 square meters)
- (Then when we visualize, we can show separate graphs for each site/habitat combination, and add a regression line with confidence intervals).


```{r}
recruitment_df <- coral_df %>% 
  group_by(site, transect, taxa, habitat, year) %>% 
  summarize(recruitment = sum(dyn_recruitment)) %>% 
  filter(transect %in% c("T01", "T02", "T03", "T04")) %>%
  mutate(recruitment_density = recruitment / 5) %>%
  ungroup()

```


## Plotting

Using our recruitment_df, we'll have 4 datapoints per year per species
Color by species
Put the year on the x axis, have ticks spaced by two

```{r}
recruitment_df %>% 
  ggplot(aes(x = factor(year), y = recruitment_density, color = taxa)) +
  geom_point() +
  facet_grid(rows = vars(site), cols = vars(substr(transect, 1, 2))) + # facet by site and habitat (BR vs OR)
  scale_x_discrete(breaks = seq(2013, 2024, by = 2)) +  # Set x-axis breaks every 2 years
  scale_color_manual(values = coral_colors) +  # Use custom coral colors
  labs(
    title = "Recruitment Density by Year, Taxa, Site, and Habitat",
    x = "Year",
    y = "Recruitment Density (recruits per square meter)"
  ) +
  theme_minimal()




# # Make a function that produces a plot like this but for one particular site
plot_recruitment_by_site <- function(site_name, taxa_name) {
  recruitment_df %>%
    filter(site == site_name) %>%
    filter(taxa == taxa_name) %>% 
    ggplot(aes(x = factor(year), y = recruitment_density)) +
    geom_point() +
    facet_wrap(~substr(habitat, 1, 2)) + # facet by habitat (BR vs OR)
    scale_x_discrete(breaks = seq(2013, 2024, by = 2)) +  # Set x-axis breaks every 2 years
    scale_color_manual(values = coral_colors) +  # Use custom coral colors
    labs(
      title = paste("Recruitment Density of", taxa_name, "at", site_name),
      x = "Year",
      y = "Recruitment Density (recruits per square meter)"
    ) +
    theme_minimal()
}
# 
# invoke function for all 4 sites, LTER1, 2, 4, 5
# and all 4 taxa: Poc, Por, Acr, Mil
plot_recruitment_by_site("LTER1", "Por")


# Regenerate the functions one more time, but add a geom_smooth() layer to show a trend line with confidence intervals
plot_recruitment_by_site_with_smooth <- function(site_name, taxa_name) {
  recruitment_df %>%
    filter(site == site_name) %>%
    filter(taxa == taxa_name) %>% 
    ggplot(aes(x = year, y = recruitment_density)) +
    geom_point() +
    geom_smooth(method = "loess", se = TRUE, color = "blue", fill = "lightblue") + # Add a smooth trend line with confidence interval
    facet_wrap(~substr(habitat, 1, 2)) + # facet by habitat (BR vs OR)
    scale_x_continuous(breaks = seq(2013, 2024, by = 2)) +  # Set x-axis breaks every 2 years
    scale_color_manual(values = coral_colors) +  # Use custom coral colors
    labs(
      title = paste("Recruitment Density of", taxa_name, "at", site_name),
      x = "Year",
      y = "Recruitment Density (recruits per square meter)"
    ) +
    theme_minimal()
}

# Test
plot_recruitment_by_site_with_smooth("LTER1", "Por")
plot_recruitment_by_site_with_smooth("LTER5", "Poc")


# plot_recruitment_by_site_with_error <- function(site_name, taxa_name) {
#   # Summarize means and SE by year and habitat
#   summary_df <- recruitment_df %>%
#     filter(site == site_name, taxa == taxa_name) %>%
#     group_by(year, habitat) %>%
#     summarise(
#       mean_recruitment = mean(recruitment_density, na.rm = TRUE),
#       se_recruitment = sd(recruitment_density, na.rm = TRUE) / sqrt(n()),
#       .groups = "drop"
#     )
#   
#   ggplot(summary_df, aes(x = year, y = mean_recruitment)) +
#     geom_point(size = 3, color = "blue") +
#     geom_errorbar(aes(ymin = mean_recruitment - se_recruitment,
#                       ymax = mean_recruitment + se_recruitment),
#                   width = 0.2, color = "blue") +
#     facet_wrap(~substr(habitat, 1, 2)) +
#     scale_x_continuous(breaks = seq(2013, 2024, by = 2)) +
#     labs(
#       title = paste("Mean Recruitment Density of", taxa_name, "at", site_name),
#       x = "Year",
#       y = "Recruitment Density (recruits per m²)"
#     ) +
#     theme_minimal()
# }
# 
# # Test this
# plot_recruitment_by_site_with_error("LTER1", "Por")


```




```{r plot recruitment v3}

# Now we're going to define a function that takes in a site and habitat input
# and then outputs a bar plot
# the x will be the year
# the y will be the mean recruit density
# we'll facet_wrap by species
# and there will be a standard error bar

# First let's try this with just a LTER5 OR
recruitment_df %>% 
  filter(site == "LTER5", habitat == "OR") %>% 
  # filter(taxa != "Mil") %>% 
  group_by(year, taxa) %>% 
  summarize(
    mean_recruitment_density = mean(recruitment_density),
    se_recruitment_density = sd(recruitment_density) / sqrt(n())
  ) %>%
  ggplot(aes(x = year, y = mean_recruitment_density, fill = taxa)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = mean_recruitment_density - se_recruitment_density,
                    ymax = mean_recruitment_density + se_recruitment_density),
                width = 0.2, position = position_dodge(0.9)) +
  facet_wrap(~taxa, ncol = 2) +
  scale_x_continuous(breaks = seq(2013, 2024, by = 2)) +  # Set x-axis breaks every 2 years
  scale_fill_manual(values = coral_colors) +
  labs(
    title = "Mean Recruitment Density at LTER5 OR",
    x = "Year",
    y = "Mean Recruitment Density (recruits per m²)"
  ) +
  theme_minimal()


```
```{r}
# Now let's make this a function
plot_recruitment_barplot <- function(site_name, habitat_name) {
  recruitment_df %>% 
    filter(site == site_name, habitat == habitat_name) %>% 
    group_by(year, taxa) %>% 
    summarize(
      mean_recruitment_density = mean(recruitment_density),
      se_recruitment_density = sd(recruitment_density) / sqrt(n())
    ) %>%
    ggplot(aes(x = year, y = mean_recruitment_density, fill = taxa)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    geom_errorbar(aes(ymin = mean_recruitment_density - se_recruitment_density,
                      ymax = mean_recruitment_density + se_recruitment_density),
                  width = 0.2, position = position_dodge(0.9)) +
    facet_wrap(~taxa, ncol = 2) +
    scale_x_continuous(breaks = seq(2013, 2024, by = 2)) +  # Set x-axis breaks every 2 years
    scale_fill_manual(values = coral_colors) +
    labs(
      title = paste("Mean Recruitment Density at", site_name, habitat_name),
      x = "Year",
      y = "Mean Recruitment Density (recruits per m²)"
    ) +
    theme_minimal()
}

# Call the function for all combos of LTER1,2,4,5 and OR, BR
plot_recruitment_barplot("LTER1", "OR")
plot_recruitment_barplot("LTER1", "BR")
plot_recruitment_barplot("LTER2", "OR")
plot_recruitment_barplot("LTER2", "BR")
plot_recruitment_barplot("LTER4", "OR")
plot_recruitment_barplot("LTER4", "BR")
plot_recruitment_barplot("LTER5", "OR")
plot_recruitment_barplot("LTER5", "BR")

# Do this in the body of a 4 loop and then save them as individual png files
# to figures/recruitment_barplots_v3_SITE_HABITAT.png
for (site in c("LTER1", "LTER2", "LTER4", "LTER5")) {
  for (habitat in c("OR", "BR")) {
    plot <- plot_recruitment_barplot(site, habitat)
    ggsave(filename = here("figures", paste0("recruitment_barplot_v3_", site, "_", habitat, ".png")),
           plot = plot,
           width = 10,
           height = 6)
  }
}

```


