

```{r}
# Clean env
rm(list = ls())

# Load necessary libraries
library(tidyverse)
library(here)
library(janitor)
library(readxl)

# Load coral data from excel file
# coral_df_raw <- read_excel(here("data", "coral_raw_wide_2023.xlsx")) %>% 
#   clean_names()
# ^ Commented out for the time-being, all our aggregated data is spread out through other sheets
# We'll need to load all the data from these individual sheets and combine them with each other
# in order to obtain a coral_df_raw

RAW_DATA_FILEPATH <- here("data_raw", "coral_raw_wide_2013-2023.xlsx")

# Get the column names of each sheet in the file that matches the pattern "LTERXAA"
sheets_habitats <- excel_sheets(RAW_DATA_FILEPATH) %>% 
  str_subset("^LTER[0-9]{1}[A-Z]{2}$")

COLUMN_LIMIT <- 100 # The maximum number of columns we'll need to load from each sheet.
                    # May need to be tweaked as dataset grows

# Load each of the sheets and join them togethr
# by stacking the datasets on top of each other.
# The columns have significant overlap, but there might be some unique ones
# among the individual sheets
# Convert each sheet to be all string type to avoid type conflicts
# coral_df_raw <- sheets_habitats %>%
#   map(~read_excel(here("data", "coral_raw_wide_2023.xlsx"), 
#                   sheet = .x, range = cell_cols(c(1,COLUMN_LIMIT)))
#       ) %>%
#   map(~mutate_all(., as.character)) %>%
#   map(~clean_names(.)) %>%
#   reduce(bind_rows)
coral_habitats_raw_cleaned_names <- sheets_habitats %>%
  map(~ read_excel(RAW_DATA_FILEPATH, 
                   sheet = .x, range = cell_cols(c(1, COLUMN_LIMIT)))
      ) %>%
  map(~ mutate_all(., as.character)) %>%
  map(~ clean_names(.)) %>%
  map(~ rename_with(., ~ gsub("_", "", .)))

coral_df_raw <- coral_habitats_raw_cleaned_names %>%
  reduce(bind_rows)

# To see where unique values are coming from, that may be weird
# coral_habitats_raw_cleaned_names[[5]] %>%
#   select("site") %>%
#   map(~unique(.x))

# View all colnames for coral_habitats_raw_cleaned_names
# filter so that they must start with 'obs'
# coral_habitats_raw_cleaned_names %>% map(~colnames(.)) %>% map(~str_subset(.x, "^obs"))
# coral_habitats_raw_cleaned_names %>% map(colnames)

# For item 3 of coral_habitats_raw_cleaned_names,
# get all distinct values of obs202263 and obs202264
# do them separately
# coral_habitats_raw_cleaned_names[[3]]$obs202263 %>% unique()
# coral_habitats_raw_cleaned_names[[3]]$obs202264 %>% unique()

  # map( # Remove underscores from column names
  #   ~rename_with(., ~str_remove(., "_"), everything())
  # ) %>%

# coral_df_raw <- sheets_habitats %>% 
#   map(~read_excel(here("data", "coral_raw_wide_2023.xlsx"), sheet = .x, range = cell_cols(c(1,90)))) %>%
#   map(~mutate_all(., as.character)) %>%
#   map(~clean_names(.)) %>% 
#   map( # Remove underscores from column names
#     ~rename_with(., ~str_remove(., "_"), everything())
#   )

# Get col names of each element in coral_df_raw
# coral_df_raw %>% map(~colnames(.))

# do again, but get the name of each sheet alongside the columns
# coral_df_raw %>% 
#   map(~colnames(.)) %>% 
#   map2(sheets_habitats, ~set_names(.x, .y))

# sheets_habitats %>% 
#   map(~read_excel(here("data", "coral_raw_wide_2023.xlsx"), sheet = .x) %>% colnames())
```

## Tidying Data

There are a lot of columns we don't need
```{r}
# First, we only want to select the columns we're interested in:
# site, habitat, transect, taxa, x, y, z, lYYYY, wYYY, hYYYY, obsYYYY, note_YYYY, note_YYYY_extra
# Make sure that for the column names with years, the exact format is followed.
# OUTDATED: We'll also add a coral_number field for each row of the raw data frame, just so we can track each coral by its distinct identity
coral_df <- coral_df_raw %>% 
  # mutate(coral_number = row_number()) %>%
  select(site, habitat, transect, taxa, x, y, z, 
         matches("l[0-9]{4}"), 
         matches("w[0-9]{4}"), 
         matches("h[0-9]{4}"), 
         matches("obs[0-9]{4}"), 
         matches("note[0-9]{4}"), 
         matches("note[0-9]{4}extra"))

# Get colnames of this new df
colnames(coral_df)


# Because it'll make things a little easier down the line, let's get the noteYYYYextra columns to instead use the format note_extraYYYY (ensuring that the col name ends in a year)
coral_df <- coral_df %>% 
  rename_with(~str_replace(., "note([0-9]{4})extra", "note_extra\\1"), matches("note[0-9]{4}extra"))

```

```{r}
###  DEBUG
# Let's spot instances of the same coral being tracked in multiple rows
# To do this we can group by site, habitat, transect, taxa, x, y, z, get the counts of each, and filter to only get the ones with counts > 1
duplicate_corals <- coral_df %>% 
  group_by(site, habitat, transect, taxa, x, y, z) %>% 
  summarise(count = n()) %>% 
  filter(count > 1) %>% 
  ungroup() %>% 
  arrange(desc(count)) %>% 
  select(-count)

# Now, let's show the rows of the original data frame that match these corals, and show the duplicate rows side by side
# loop through the coral_df rows, and check if any of them match a row inside duplicate_corals
# initialize an object to track collected duplicate rows
duplicate_rows_collection <- data.frame()
for(coral in 1:nrow(coral_df)){
  # Get the row of coral_df
  coral_row <- coral_df[coral,]
  # Get the row of duplicate_corals
  coral_row_duplicate <- duplicate_corals %>% 
    filter(site == coral_row$site,
           habitat == coral_row$habitat,
           transect == coral_row$transect,
           taxa == coral_row$taxa,
           x == coral_row$x,
           y == coral_row$y,
           z == coral_row$z)
  # If the row exists, add the rows to the duplicate_rows_collection
  if(nrow(coral_row_duplicate) > 0){
    # Add the row to the collection
    duplicate_rows_collection <- bind_rows(duplicate_rows_collection, coral_row)
  }
}

# write to file system
# write_csv(duplicate_rows_collection, here("data_outputs", "DEBUG_coral_duplicates.csv"))

```

```{r}
```



There are two note columns the regual one and the 'extra' one.

Let's tidy this data. We'll want to have a year column. We'll also want to have a column for each of length, width, height, observer, and note. 

```{r}

# First convert anything numeric to be a string to avoid type conflicts
# coral_df <- coral_df %>% 
#   mutate(across(where(is.numeric), as.character))

coral_df_tidy <- coral_df %>%
  pivot_longer(cols = matches("^l\\d{4}$|^w\\d{4}$|^h\\d{4}$|^obs\\d{4}$|^note\\d{4}$|^note_extra\\d{4}$"), 
               names_to = "key", 
               values_to = "value") %>%
  mutate(year = str_extract(key, "\\d{4}"))

# Trim off the year from every thing in the key column (including the underscore that sometimes precedes dates)
coral_df_tidy <- coral_df_tidy %>% 
  mutate(key = str_remove(key, "_*\\d{4}"))

# Now pivot wider to get the columns we want
coral_df_tidy <- coral_df_tidy %>%
  pivot_wider(names_from = key, values_from = value) %>%
  rename(length = l, width = w, height = h, observer = obs)
```
```{r}
### DEBUG: We're getting List Columns! Let's check it out.
# Let's get a comprehensive list of all our list columns
# Grab all that are LTER4_OR_T03_Acr_0.75_0_0
# coral_df_tidy %>% filter(
#   site == "LTER4",
#   habitat == "OR",
#   transect == "T03",
#   taxa == "Acr",
#   x == 0.75,
#   y == 0,
#   z == 0
# )
# YES CONFIRMED, we have list columns,
# let's get them all so we can see exactly which is which

# list_cols <- coral_df_tidy %>% 
#   select(where(~is.list(.x))) %>% 
#   colnames()

# This didn't work, let's try another method
# Get all the columns that are list columns
# list_cols <- sapply(coral_df_tidy, is.list)
# names(coral_df_tidy)
```


So we aren't getting any list-cols! That's good.

## Re-Converting numeric rows and removing NA values (Discontinued)

Now the last thing we need to do is convert the fields x, y, z, year, length, width, height to be properly numeric again.

```{r}
# coral_df_tidy <- coral_df_tidy %>%
#   mutate_at(vars(x, y, z, year, length, width, height), as.numeric)
# NAs will be introduced by this for string values, and that's ok.
```

```{r}
# This commmented code below turns all "NA", "Na", and "UK" values to be NA
# We will not do this, as per Mohsen's request
# coral_df_tidy <- coral_df_tidy %>%
#   mutate_at(vars(observer, note, note_extra), ~na_if(., "NA")) %>%
#   mutate_at(vars(observer, note, note_extra), ~na_if(., "Na")) %>%
#   mutate_at(vars(observer, note, note_extra), ~na_if(., "UK"))
```

## Normalizing Factor Variables

There are some discrete factor variables that may have different cases.
- habitat should be set to all uppercase
- taxa should be set to title case
- observer should be set to upper case
- site should be uppercase
- transect should be uppercase

```{r}
coral_df_tidy <- coral_df_tidy %>%
  mutate(habitat = toupper(habitat),
         taxa = str_to_title(taxa),
         observer = toupper(observer),
         site = toupper(site),
         transect = toupper(transect))
```


```{r}
## Code to check the result of normalization
# Show distinct values for each non numeric field individually
# coral_df_tidy %>% 
#   select(habitat, taxa, observer, site, transect) %>% 
#   map(~unique(.x))
```
## Normalize the NA string values

If a capital string NA is written, R gets confused when reading and thinks it's truly an NA, when I want to keep it as an NA string. To get around this, I'll make sure that all na strings are written as Na (title case).

```{r}
# For all cells, if the value is "NA", convert it to "Na"
coral_df_tidy <- coral_df_tidy %>%
  mutate_all(~str_replace(., "NA", "Na"))

```



## Renaming Transects and Plots

Right now P10 and P1 get placed right next to each other because it's alphabetical. I want to rename any one digit transects to have a leading zero.

```{r}
coral_df_tidy_transect_renamed <- coral_df_tidy %>%
  mutate(transect = str_replace(transect, "^([PT])([0-9])(_.*)?$", "\\10\\2\\3"))

# coral_df_tidy_transect_renamed %>% 
#   select(habitat, taxa, observer, site, transect) %>% 
#   map(~unique(.x))
```

## Storing

Let's store this data file so that we can use it later.

```{r}
write_csv(coral_df_tidy_transect_renamed, here("data_outputs", "coral_tidy_2013-2023.csv"))
```



