
In this file, we're going to load the tidy data created in tidy_coral_data.Rmd and extract information from the note columns to populate new dynamic columns.

The dynamic columns we'll make are death, recruitment, growth, fission, fusion, and f-group.

## Setup
```{r}
# Clean workspace
rm(list = ls())

# Load necessary libraries
library(tidyverse)
library(here)
library(janitor)
library(tsibble)

# TEST param. Set to True if you want to work with a truncated version of the dataset
TEST <- FALSE

# Load our tidy coral dataset
# if we're testing, we'll load the test one. coral_tidy_2024_for_test.csv
if(TEST) {
  coral_df_tidy <- read_csv(here("data_outputs", "coral_tidy_2024_for_test2.csv"))
} else {
  coral_df_tidy <- read_csv(here("data_outputs", "coral_tidy_2024.csv"))
}

# We want to convert this to a tibble to do some time series analysis, and a
# key is needed to distinguish distinct corals from each other.
# thus we can create a coral_number field that will uniquely identify corals
# we'll obtain this by stringing together the site, habitat, transect, taxa, x, y, z, separating by underscores
coral_df_tidy <- coral_df_tidy %>% 
  mutate(coral_number = paste0(site, "_", habitat, "_", transect, "_", 
                               taxa, "_", x, "_", y, "_", z))



# Convert to tsibble
coral_ts <- coral_df_tidy %>%
  as_tsibble(index = year, key=coral_number)
# ^ For some reason this isn't working

```

```{r}
### Catch the Duplicates!
glimpse(coral_df_tidy)
duplicates(coral_df_tidy, key=coral_number, index=year)

# duplicates <- coral_df_tidy %>%
#   count(coral_number, year) %>%
#   filter(n > 1)
# 
# # Filter for the coral_numbers being equal to
# # LTER4_OR_T03_Acr_0.75_0_0

# lter4_or_t03_acr_0.75_0_0 <- coral_df_tidy %>%
#   filter(coral_number == "LTER4_OR_T03_Acr_0.75_0_0") %>%
#   select(year, length, width, height, note, note_extra)

```




## Aggregating and tokenizing notes

```{r}
# Let's extract the note values for each observation and place them in a list in each column
# we'll have a temporary field called notes_extracted that will hold the individual tokenized notes as a list

# To extract the notes, take the column note and note_extra, splice each along commas and semicolons, trim, and convert to lowercase. Combine the lists from note and note_extra to get the value for notes_extracted


# coral_ts <- coral_ts %>% 
#   mutate(notes_extracted = map2(note, note_extra, ~c(strsplit(tolower(.x), "[,;]")[[1]], strsplit(tolower(.y), "[,;]")[[1]])))

coral_ts <- coral_ts %>% 
  mutate(notes_extracted = map2(note, note_extra, ~c(strsplit(tolower(.x), "[,;]")[[1]], strsplit(tolower(.y), "[,;]")[[1]]))) %>% 
  mutate(notes_extracted = map(notes_extracted, ~str_trim(.x)))
```

```{r}
# TEMP: Exploration of unique values in notes.
# Get all the unique values of notes_extracted
all_notes <- coral_ts %>%
  filter(year >= 2019) %>%
  pull(notes_extracted) %>%
  unlist() %>%
  unique()

all_notes
rm(all_notes)
```

## Setting up first_instance & coral_encountered properties

Let's initialize a property called first_instance that will indicate
if data observed represents the first instance of a particular coral
appearing in the dataset. This will be helpful later on.

Strategy:
- Create a filtered version of the coral_ts that excludes any observations that have non-numeric data for the dimensions of the coral.
- In this new filtered dataset, get the min(year) for each unique coral number, and set a new vector called first_instance_years vector to these values.
- Then modify the original coral_ts that by setting first_instance=TRUE if the year of the observation matches the corresponding year for that coral in first_instance_years

```{r}
coral_ts_numeric_only <- coral_ts %>% 
  filter(!is.na(suppressWarnings(as.numeric(length))) & 
           !is.na(suppressWarnings(as.numeric(width))) & 
           !is.na(suppressWarnings(as.numeric(height))))

# Get the first instance year for each coral
# We want to get it as a list, so that way we can map coral_number to a given year that it was first encountered
first_instance_years <- coral_ts_numeric_only %>% 
  as_tibble() %>%
  group_by(coral_number) %>% 
  summarize(first_instance_year = min(year)) %>% 
  deframe() %>% 
  as.list()

# Note to access a year value here, the coral number needs to be turned into a string and then used as a key.


# first_instance_years <- coral_ts_numeric_only %>% 
#   as_tibble() %>%
#   group_by(coral_number) %>% 
#   summarize(first_instance_year = min(year)) %>% 
#   arrange(coral_number) %>% 
#   pull(first_instance_year)

# coral_ts <- coral_ts %>%
#   mutate(first_instance = year == first_instance_years)

# initialize empty columns
coral_ts <- coral_ts %>% 
  mutate(first_instance = FALSE, coral_encountered = FALSE)

# Loop through the coral_ts and set first_instance to TRUE if the year matches the first_instance_years and coral_encountered to TRUE if the year is greater than or equal to the first_instance_years
for(i in 1:nrow(coral_ts)) {
  # Some corals don't have a first instance year, skip these
  first_instance_year <- first_instance_years[[coral_ts$coral_number[i] %>% as.character()]]
  if(is.null(first_instance_year)) {
    next
  }
  coral_ts$first_instance[i] <- coral_ts$year[i] == first_instance_year
  coral_ts$coral_encountered[i] <- coral_ts$year[i] >= first_instance_year
}


```

TODO move explanation above?
Next let's use the first_instance property to create yet another property called coral_encountered. This will be set to TRUE if the coral has been encountered before, and FALSE otherwise.


```{r}
# coral_ts <- coral_ts %>%
#   mutate(coral_encountered = year >= first_instance_years)
```


## Parsing notes for dynamics

### Death
```{r}
# Initialize empty death column
coral_ts <- coral_ts %>% mutate(death = 0)

# If notes_extracted contains a "d", then set the new column "death" to 1. 
# It should be 0 otherwise.
for (i in 1:nrow(coral_ts)) {
  note <- coral_ts$notes_extracted[[i]]
  if ("d" %in% note || "dead" %in% note) {
    coral_ts$death[i] <- 1
  }
  # There are some corals where death was not indicated in the note,
  # but in the dimension fields by writing "Dead" or "D" for length, width, and height.
  # Let's just check if the length field lowercase equals d or death
  length <- coral_ts$length[i] %>% tolower()
  if(!is.na(length) && (length == "d" || length == "dead")) {
    coral_ts$death[i] <- 1
  }
}
```

#### Death Years and coral_died property
It'll be useful to get these things initialized. It will help us to avoid marking something with the death dynamic multiple years, as well as help us with identifying missing data.

First, let's create  death_years by doing filter for death_dynamic and group_by for coral_numbers, then do a min() for each group to get the death year.
```{r}
death_years <- coral_ts %>% 
  as_tibble() %>% 
  filter(death==1) %>% 
  group_by(coral_number) %>% 
  summarize(death_years = min(year)) %>% 
  deframe() %>% 
  as.list()
```

Then let's create the field coral_died by using a for loop to compare each observation's year with the death_year. If it's greater than or equal to it's corresponding death year,
we say it died.

```{r}
# Initialize empty col
coral_ts <- coral_ts %>% 
  mutate(coral_died = FALSE)

# Set the property
for(i in 1:nrow(coral_ts)) {
  death_year <- death_years[[coral_ts$coral_number[i] %>% as.character()]]
  # Leave ones that haven't died yet (ie no death year) to remain as FALSE
  if(is.null(death_year)) {
    next
  }
  coral_ts$coral_died[i] <- coral_ts$year[i] >= death_year
}
```



### Recruitment (Explicitly Noted)
```{r}
# First initialize an empty recruitment column
coral_ts <- coral_ts %>% mutate(recruitment = 0)

# If notes_extracted contains "r", "(r)", "r-fr", "recruit" or "recruitment..."  then set the  new column 
# recruitment to 1. It should be 0 otherwise
for (i in 1:nrow(coral_ts)) {
  notes <- coral_ts$notes_extracted[[i]]
  for(note in notes) {
    if(is.na(note)) {
      next
    }
    if(note == "r" ||
       note == "(r)" ||
       note == "r-fr" ||
       note == "recruit" ||
       grepl("^recruitment", note)) {
      coral_ts$recruitment[i] <- 1
    }
  }
}

```

### Recruitment (Implicit) and Missing Data
There are some cases where a never before documented coral suddenly appears.
For the sake of this study we will consider any of these corals that suddenly appear as recruits if their largest dimension is less than 6cm. Otherwise, we will need to label this observation with the dynamic "missing data".

**NOTE: The following rule no longer applies as of Oct 2024**
There is another case for missing data that we need to handle: any time there is a gap for a particular coral. We can tell if this is the case because first_instance is FALSE and there is no numerical data for the dimensions of the coral in the previous year.

**This rule was disabled because** Hunter and Mohsen decided that they definitely would not have missed any corals while doing the survey. It must be that at some point they just forgot to note "Dead" in some way and decided that a lack of numerical data was enough to imply death.


```{r}
# First, let's initialize the missing_data column.
# (We don't need to re-initialize the recruitment column)
coral_ts <- coral_ts %>% mutate(missing_data = 0)

# Loop through the coral_ts
# Check if first_instance=TRUE
#   if the largest dimension is less than 6, set recruitment to 1
#.  otherwise, set missing_data to 1
for(i in 1:nrow(coral_ts)) {
   # First get all the coral dimensions with as.numeric, and suppress warnings
  coral_length <- suppressWarnings(as.numeric(coral_ts$length[i]))
  coral_width <- suppressWarnings(as.numeric(coral_ts$width[i]))
  coral_height <- suppressWarnings(as.numeric(coral_ts$height[i]))
    
  if(coral_ts$first_instance[i] == TRUE) {
    # Get the largest dimension
    # remember that length, width, height are str type.
    # Sometimes they might say "Na" or "Dead" instead of a number.
    # These should be skipped. Otherwise, convert to a number first.
    if(is.na(coral_length) || is.na(coral_width) || is.na(coral_height)) {
      next
    }
    largest_dim <- max(coral_length, coral_width, coral_height)
    if(largest_dim < 6) {
      coral_ts$recruitment[i] <- 1
    } else if(coral_ts$year[i] > min(coral_ts$year)) {
      coral_ts$missing_data[i] <- 1
    }
  } 

}

```


## Explicitly Noted MissingData

For cases where the notes contain "unsampled" or "not sampled", we will set the missing_data dynamic to 1.

```{r}
# Loop through coral_ts
for(i in 1:nrow(coral_ts)) {
  notes <- coral_ts$notes_extracted[[i]]
  
  for (note in notes) {
    if(is.na(note)) {
      next
    }
    if(note == "unsampled" || note == "not sampled") {
      coral_ts$missing_data[i] <- 1
    }
  }
}
```


### Vanishing Corals (Another Type of Death)

There are some cases where a coral that was previously documented is no longer present. We will consider these corals as having died, and will set the death dynamic to 1 for these observations.

Exception: if a coral is explicitly noted as "unsampled" or "not sampled", then the death dynamic won't be applied.

```{r}
# Vanishing Logic
# If the coral suddenly vanished
# (meaning there is no numerical data for its dimensions, 
# but it hasn't been confirmed dead in some other way yet), then 
# call it dead and update the other vars as necessary.

# We'll need to loop through in order of the years for this to work correctly
coral_ts <- coral_ts %>% arrange(year)

# Make a vector of vanished corals.
# We'll grow this as we loop through and add vanished corals we find here, that way we don't double mark any as dead.
vanished_corals <- c()
  
# Loop through coral ts
for(i in 1:nrow(coral_ts)) {
  # First check if the current coral number observation has already been found so far. Skip if so.
  if(coral_ts$coral_number[i] %in% vanished_corals) {
    next
  }
  
  # If the year is 2013 (the first year of the study), then we'll also skip because no coral can vanish in the first year.
  if(coral_ts$year[i] == 2013) {
    next
  }
  
  # get all the coral dimensions with as.numeric, 
  # and suppress warnings
  coral_length <- suppressWarnings(as.numeric(coral_ts$length[i]))
  coral_width <- suppressWarnings(as.numeric(coral_ts$width[i]))
  coral_height <- suppressWarnings(as.numeric(coral_ts$height[i]))
  
  # Check if the notes contain "unsampled" or "not sampled"
  noted_as_unsampled <- FALSE
  notes <- coral_ts$notes_extracted[[i]]
  for (note in notes) {
    if(is.na(note)) {
      next
    }
    if(note == "unsampled" || note == "not sampled") {
      noted_as_unsampled <- TRUE
      break
    }
  }
  
  # Next if the coral has been encountered, has NOT been 
  # confirmed dead already via the comments or LWH values, 
  # AND if there is no current numerical data, 
  # AND if it hasn't been explicitly noted as 'not sampled' or 'unsampled',
  # it's safe to say 
  # that it died that year.
  if(coral_ts$coral_encountered[i] == TRUE &&
       coral_ts$coral_died[i] == FALSE &&
       is.na(coral_length) && 
       is.na(coral_width) && 
       is.na(coral_height) &&
       noted_as_unsampled == FALSE) {
    # browser()
    coral_ts$death[i] <- 1
    # Update our vanished_coral vec
    vanished_corals <- c(vanished_corals, coral_ts$coral_number[i])
  }
}
    

```

### Fusion
We know that fusion ocurred at an observation
- if the token itself EQUALS 'fuX', 'fusX', 'fuX-ed' where X consists of zero or more digits
- if the token itself contains the word "fused" or "fusion"

In the first case, the string of digits X should be saved in a field called fusion_group


```{r}
# Start by initializing empty columns fusion and fusion_group
coral_ts <- coral_ts %>% 
  mutate(fusion = 0, fusion_group = NA)

for(i in 1:nrow(coral_ts)) {
  notes <- coral_ts$notes_extracted[[i]]
  
  # Iterate through all values in notes
  for (note in notes) {
    # print(paste("Note is ", note))
    if (grepl("^fu[0-9]*$", note) || 
        grepl("^fus[0-9]*$", note) || 
        grepl("^fu[0-9]*-ed$", note)) {
      # This is the ideal case where we have a fuX or fusX or fuX-ed.
      # We'll be able to extract a fusion_group number from this format
      coral_ts$fusion[i] <- 1
      fus_group <- str_extract(note, "^fus?([0-9]*)(-ed)?$", group = 1)
      coral_ts$fusion_group[i] <- as.numeric(fus_group)
    } else if(grepl("fused", note) || grepl("fusion", note)) {
      # It's possible that we'll get these sorts of notes, but we shouldn't
      # overwrite any existing groups, that's why if fusion has already been set
      # we'll just leave it as is.
      if(coral_ts$fusion[i] == 1) {
        next
      }
      coral_ts$fusion[i] <- 1
      coral_ts$fusion_group[i] <- NA
    }
  }
}

```


### Fission
This is similar to the fusion case, but with less possibilities
We label fission as having ocurred if
- if the token itself EQUALS 'fiX' or 'fisX' where consists of zero or more digits
- if the token contains the word 'fission'

```{r}
# Start by initializing empty columns fission and fission_group
coral_ts <- coral_ts %>% 
  mutate(fission = 0, fission_group = NA)

for(i in 1:nrow(coral_ts)) {
  notes <- coral_ts$notes_extracted[[i]]
  
  # Iterate through all values in notes
  for (note in notes) {
    if (grepl("^fis?[0-9]*$", note)) {
      # This is the ideal case where we have a fiX or fisX.
      # We'll be able to extract a fission_group number from this format
      coral_ts$fission[i] <- 1
      fis_group <- str_extract(note, "^fis?([0-9]*)$", group = 1)
      coral_ts$fission_group[i] <- as.numeric(fis_group)
    } else if(grepl("fission", note)) {
      # It's possible that we'll get these sorts of notes, but we shouldn't
      # overwrite any existing groups, that's why if fission has already been set
      # we'll just leave it as is.
      if(coral_ts$fission[i] == 1) {
        next
      }
      
      
      
      coral_ts$fission[i] <- 1
      coral_ts$fission_group[i] <- NA
    }
  }
}


```

### Growth & Shrinkage
here we won't actually want to use the comments field, but rather the growth ratios, which we will need to compute
```{r}
# Compute growth_ratio field for each observation
coral_ts <- coral_ts %>% 
  group_by(coral_number) %>%
  mutate(
    sum_dims = as.numeric(length) + as.numeric(width) + as.numeric(height),  # Sum of current year's dimensions
    prev_sum_dims = lag(sum_dims),  # Sum of previous year's dimensions
    growth_ratio = sum_dims / prev_sum_dims  # Calculate the growth ratio
  ) %>% 
  select(-sum_dims, -prev_sum_dims) %>%
  ungroup() 

# Set the fields for growth and shrinkage
# growth should be 1 if growth_ratio is >1, 0 otherwise
# shrinkage should be 1 if growth_ratio is <1, 0 otherwise
coral_ts <- coral_ts %>% 
  mutate(growth = 0, shrinkage = 0)

for(i in 1:nrow(coral_ts)) {
  # First make sure that growth_ratio is not NA. if it is, then next
  if(is.na(coral_ts$growth_ratio[i])) {
    next
  }
  if (coral_ts$growth_ratio[i] > 1) {
    coral_ts$growth[i] <- 1
  } else if (coral_ts$growth_ratio[i] < 1) {
    coral_ts$shrinkage[i] <- 1
  }
}

```

### Immigration & Emigration
```{r}
# Set immigration to 1 if notes_extracted starts with "im", 0 otherwise
# Set emigration to 1 if notes_extracted starts with "em", 0 otherwise
coral_ts <- coral_ts %>% 
  mutate(immigration = 0, emigration = 0)

for(i in 1:nrow(coral_ts)) {
  notes <- coral_ts$notes_extracted[[i]]
  
  # Iterate through all values in notes
  for (note in notes) {
    if (grepl("^im", note)) {
      coral_ts$immigration[i] <- 1
    }
    if(grepl("^em", note)) {
      coral_ts$emigration[i] <- 1
    }
  }
}
```

### Edge
```{r}
# Set edge to 1 if a note contains "edge", is equal to "(fu-)?ed",  0 otherwise
coral_ts <- coral_ts %>% 
  mutate(edge = 0)

for(i in 1:nrow(coral_ts)) {
  notes <- coral_ts$notes_extracted[[i]]
  
  # Iterate through all values in notes
  for (note in notes) {
    if (grepl("edge", note) || grepl("^(fu[0-9]*-)?ed$", note)) {
      coral_ts$edge[i] <- 1
    }
  }
}
```


## Extra: Exploration

This code block was used to explore the distinct values of notes for the purposes of parsing them into dynamics.
```{r}
# First let's explore the notes to see the types of values that exist
# all_notes <- coral_ts %>% 
#   pull(notes_extracted) %>% 
#   unlist() %>% 
#   unique()
# 
# # In all_notes, what values start with "em" or "im"
# all_notes %>%
#    str_subset("edge")


# filter out NAs
# %>%
#   filter(str_detect(., "immig|emig")) %>%

```

## Naming dynamics fields
Because there are a lot of dynamics field, we'll want to rename them to preface them with dyn_ to make them easier to identify

```{r}
# Rename all dynamic fields to start with "dyn_"
coral_ts <- coral_ts %>% 
  rename(dyn_growth = growth,
         dyn_shrinkage = shrinkage,
         dyn_death = death,
         dyn_recruitment = recruitment,
         dyn_fission = fission,
         dyn_fusion = fusion,
         dyn_edge = edge,
         dyn_immigration = immigration,
         dyn_emigration = emigration,
         dyn_missing_data = missing_data)
```

## Storage
```{r}
# remove the notes_extracted column, then save it as a csv
# let's call it "coral_tidy_2024_with_dynamics.csv"
coral_ts <- coral_ts %>% select(-notes_extracted, -first_instance, -coral_encountered, -coral_died) %>% 
  write_csv(here("data_outputs", "coral_tidy_2024_with_dynamics.csv"))
```




