
This file will be in charge of doing all the work needed to convert the widened cleaned data into a nice excel with all the needed bells and whistles that make viewing the results easier.

This file will be in charge of adding conditional formatting rules (highlights)
to the widened data. This is for the purpose of making the data easier to read.

```{r}
# Clean env
rm(list=ls())

# Load necessary libraries
library(tidyverse)
library(here)
library(openxlsx)

# Load data
coral_clean_wide <- read_csv(here("data_outputs", "coral_clean_wide_2024.csv"))
```

## Col Name Changes
There are a set of names that should change to be like what was in the original spreadsheet.
These names are more readable, and it'll make the scripts be backwards compatible with the produced human-friendly excel files; the outputed file can be edited and RE-fed into the sequence of scripts to be re-cleaned.
```{r}
# First, rename height_XXXX, widith_XXXX, and length_XXXX to just hXXXX, wXXXX, and lXXXX
# also rename observer_XXXX to just obsXXXX.
# also rename "note_extra_XXXX" to "note_XXXX_extra"


coral_clean_wide <- coral_clean_wide %>% 
  rename_with(~str_replace(., "height_", "H"), starts_with("height")) %>% 
  rename_with(~str_replace(., "width_", "W"), starts_with("width")) %>% 
  rename_with(~str_replace(., "length_", "L"), starts_with("length")) %>% 
  rename_with(~str_replace(., "observer_", "Obs"), starts_with("observer")) %>% 
  rename_with(~ str_replace(., "note_extra_(\\d+)", "note_\\1_extra"))


```



## Create workbook and add sheets
```{r}
wb <- createWorkbook()

# First will have a sheet that shows all the data collected in the study
addWorksheet(wb, "ALL_SITES")
writeData(wb, sheet = "ALL_SITES", x = coral_clean_wide)

# Then we'll have individual sheets for each site and habitat
site_habitat <- coral_clean_wide %>% select(site, habitat) %>% distinct()
for(i in 1:nrow(site_habitat)){
  site <- site_habitat$site[[i]]
  habitat <- site_habitat$habitat[[i]]
  sheet_name <- paste0(site, habitat)
  addWorksheet(wb, sheet_name)
  writeData(wb, sheet = sheet_name, x = coral_clean_wide %>% filter(site == site_habitat$site[[i]], habitat == site_habitat$habitat[[i]]))
}

# Save the excel file (needed for collecting all the sheet names)
saveWorkbook(wb, file = here("data_outputs", "coral_clean_wide_human-friendly.xlsx"), overwrite = TRUE)

# Get all the sheetnames as a vector
sheet_names <- getSheetNames(here("data_outputs", "coral_clean_wide_human-friendly.xlsx"))


```

## Conditional Formatting

The first set of conditional formatting will be for the dynamics fields.
```{r}
# Create a list of style objects for dynamics:
# The background colors should be
# - Recruitment: Blue
# - Growth: Green
# - Death: Red
styles <- list(
  Recruitment = createStyle(fontColour = "black", bgFill = "#159FEC"),
  Growth = createStyle(fontColour = "black", bgFill = "#17A43F"),
  Shrinkage = createStyle(fontColour = "white", bgFill = "darkgreen"),
  Death = createStyle(fontColour = "black", bgFill = "#FB0006"),
  Fission = createStyle(fontColour = "black", bgFill = "#A1BCD6"),
  Fusion = createStyle(fontColour = "black", bgFill = "#B07EDD"),
  MultipleDynamics = createStyle(fontColour = "white", bgFill = "black"),
  NaValue = createStyle(fontColour = "black", bgFill = "yellow"),
  UkValue = createStyle(fontColour = "black", bgFill = "orange"),
  MissingData = createStyle(fontColour = "white", bgFill = "brown")
)




```

```{r}
# Apply conditional formatting for dynamics
# First try doing it just for Recruitment in ALL_SITES tab

# dyn_col_ind <- which(grepl("^dynamics", colnames(coral_clean_wide)))
# row_range <- 2:(nrow(coral_clean_wide)+1)

# # The easy ones first
# conditionalFormatting(wb, sheet = "ALL_SITES", cols = dyn_col_ind, rows = row_range, rule = '=="Recruitment"', style = styles$Recruitment)
# conditionalFormatting(wb, sheet = "ALL_SITES", cols = dyn_col_ind, rows = row_range, rule = '=="Growth"', style = styles$Growth)
# conditionalFormatting(wb, sheet = "ALL_SITES", cols = dyn_col_ind, rows = row_range, rule = '=="Death"', style = styles$Death)
# 
# # For fission and fusion, we'll use the beginsWith type
# conditionalFormatting(wb, sheet = "ALL_SITES", cols = dyn_col_ind, rows = row_range, type="beginsWith", rule = 'Fission', style = styles$Fission)
# conditionalFormatting(wb, sheet = "ALL_SITES", cols = dyn_col_ind, rows = row_range, type="beginsWith", rule = 'Fusion', style = styles$Fusion)
# 
# # For multiple dynamics, we'll use the contains type. The rule will be ","
# conditionalFormatting(wb, sheet = "ALL_SITES", cols = dyn_col_ind, rows = row_range, type="contains", rule = ",", style = styles$MultipleDynamics)
# 



```

I want to be able to do what was done for the ALL_SITES sheet, except for all the sheets in the WB

```{r}
dyn_col_ind <- which(grepl("^dynamics", colnames(coral_clean_wide)))
row_range <- 2:(nrow(coral_clean_wide)+1)

for (sheet in sheet_names){
  # easy ones first
  conditionalFormatting(wb, sheet = sheet, cols = dyn_col_ind, rows = row_range, rule = '=="Recruitment"', style = styles$Recruitment)
  conditionalFormatting(wb, sheet = sheet, cols = dyn_col_ind, rows = row_range, rule = '=="Growth"', style = styles$Growth)
  conditionalFormatting(wb, sheet = sheet, cols = dyn_col_ind, rows = row_range, rule = '=="Shrinkage"', style = styles$Shrinkage)
  conditionalFormatting(wb, sheet = sheet, cols = dyn_col_ind, rows = row_range, rule = '=="Death"', style = styles$Death)
  conditionalFormatting(wb, sheet = sheet, cols = dyn_col_ind, rows = row_range, rule = '=="Missing Data"', style = styles$MissingData)
  
  # For fission and fusion, we'll use the beginsWith type
  conditionalFormatting(wb, sheet = sheet, cols = dyn_col_ind, rows = row_range, type="beginsWith", rule = 'Fission', style = styles$Fission)
  conditionalFormatting(wb, sheet = sheet, cols = dyn_col_ind, rows = row_range, type="beginsWith", rule = 'Fusion', style = styles$Fusion)
  
  # For multiple dynamics, we'll use the contains type. The rule will be ","
  conditionalFormatting(wb, sheet = sheet, cols = dyn_col_ind, rows = row_range, type="contains", rule = ",", style = styles$MultipleDynamics)
  
  # For the Na and UK values, we'll use the default type, and apply it to all columns
  conditionalFormatting(wb, sheet = sheet, cols = 2:ncol(coral_clean_wide), rows = row_range, rule = '=="Na"', style = styles$NaValue)
  conditionalFormatting(wb, sheet = sheet, cols = 2:ncol(coral_clean_wide), rows = row_range, rule = '=="UK"', style = styles$UkValue)
  
}
```


## Final Renaming
TODO make apply to workbook.... somehow
To make this easy to read, a set of shortened names will be used for columns
```{r}
# First, rename height_XXXX, widith_XXXX, and length_XXXX to just hXXXX, wXXXX, and lXXXX
# also rename observer_XXXX to just obsXXXX.
# also rename dynamics_XXXX to just dynXXXX.
# also rename growth_ratio_XXXX to just grXXXX.
# coral_clean_wide <- coral_clean_wide %>% 
#   rename_with(~str_replace(., "height_", "h"), starts_with("height")) %>% 
#   rename_with(~str_replace(., "width_", "w"), starts_with("width")) %>% 
#   rename_with(~str_replace(., "length_", "l"), starts_with("length")) %>% 
#   rename_with(~str_replace(., "observer_", "obs"), starts_with("observer")) %>% 
#   rename_with(~str_replace(., "dynamics_", "dyn"), starts_with("dynamics")) %>% 
#   rename_with(~str_replace(., "growth_ratio_", "gr"), starts_with("growth_ratio"))
# 
# # Next, convert all names to title case
# colnames(coral_clean_wide) <- colnames(coral_clean_wide) %>% str_to_title()

```



## Save the Excel Sheet
```{r}
saveWorkbook(wb, file = here("data_outputs", "coral_clean_wide_human-friendly.xlsx"), overwrite = TRUE)
```



