```{r}
rm(list = ls())

library(tidyverse)
library(here)
library(janitor)
library(readxl)
library(openxlsx)


# Load the dataset in data folder: coral_raw_wide_2024.xlsx

coral_df <- read_excel(here("data", "coral_raw_wide_2024.xlsx")) %>% 
  clean_names()

coral_df_long_2019 <- read_csv(here("data", "coral_raw_long_2019.csv")) %>% 
  clean_names()

# LTER1BR_Dyn_finishedX <- read_excel(here("data", "LTER1BR_Dyn_finishedX.xlsx")) %>% 
#   clean_names()
# 
# LTER1OR_Dyn_finishedX <- read_excel(here("data", "LTER1OR_Dyn_finishedX.xlsx")) %>% 
#   clean_names()


```

TESTING EXCEL WRITING 
```{r}
# Create sample dataframes
df1 <- data.frame(
  Name = c("John", "Jane", "Doe"),
  Age = c(30, 25, 35),
  Occupation = c("Engineer", "Doctor", "Artist")
)

df2 <- data.frame(
  Product = c("Apple", "Banana", "Cherry"),
  Price = c(1.20, 0.50, 2.00),
  Quantity = c(10, 20, 15)
)

df3 <- data.frame(
  Country = c("USA", "Canada", "Mexico"),
  Capital = c("Washington", "Ottawa", "Mexico City"),
  Population = c(331, 38, 126)
)

# Create a new workbook
wb <- createWorkbook()

# Add sheets to the workbook
addWorksheet(wb, "Sheet1")
addWorksheet(wb, "Sheet2")
addWorksheet(wb, "Sheet3")

# Write dataframes to the respective sheets
writeData(wb, sheet = "Sheet1", x = df1)
writeData(wb, sheet = "Sheet2", x = df2)
writeData(wb, sheet = "Sheet3", x = df3)


# Add a conditional highlighting rule
# addConditionalFormatting(wb, sheet = "Sheet1", header = TRUE, data = df1,
#                          rule = ConditionalFormattingRule$new("Expression", style = createStyle(fontColour = "#FF0000", halign = "center")))

# Create a style for highlighting
highlightStyle <- createStyle(
  fontColour = "#FFFFFF",        # Font color (white)
  bgFill = "#FF0000",            # Background color (red)
  halign = "CENTER",             # Horizontal alignment (center)
  valign = "CENTER"              # Vertical alignment (center)
)

# Apply the style to specific cells
# Highlight the 'Score' column for scores greater than 90
conditionalFormatting(wb, "Sheet1", cols = 2, rows = 2:4,
                      rule = "==John", style = highlightStyle)

# Save the workbook
saveWorkbook(wb, file = here("multiple_sheets.xlsx"), overwrite = TRUE)

```


I want to do a quick check on the column names of all the sheets in the file coral_raw_wide_2024.xlsx
```{r}
# Get the column names of each sheet in the file
sheets <- excel_sheets(here("data", "coral_raw_wide_2024.xlsx"))
# Get only the sheets that follow the regex pattern of "^LTERXYY$"
# but X should be a digit, and YY should be 2 letters
sheets_habitats <- sheets %>% 
  str_subset("^LTER[0-9]{1}[A-Z]{2}$")

# Get the column names of each sheet
sheets_habitats %>% 
  map(~read_excel(here("data", "coral_raw_wide_2024.xlsx"), sheet = .x) %>% colnames())

# Check if they are identical
sheets_habitats %>% 
  map(~read_excel(here("data", "coral_raw_wide_2024.xlsx"), sheet = .x) %>% colnames()) %>% 
  reduce(intersect)

```


Eventually I may wish to use length and width data along with positions to guess which corals have fissioned or fused into which new corals. (First off should understand from Hunter and Mohsen if this is important or interesting)
```{r}
# Make a plot showing relationship between width and length fields in coral_df_long_2019
# also graph a y = x line
ggplot(coral_df_long_2019, aes(x = width, y = length, color = taxa)) +
  geom_point(size = 0.5, alpha = 0.5) +
  labs(title = "Relationship between width and length of corals",
       x = "Width",
       y = "Length") +
  geom_abline(intercept = 0, slope = 1)
```

The note fields can be quite useful
```{r}
# For coral_df, get all the column names that begin with "note". 
# Get the distinct values across all of these columns.
# make all lowercase.
# remove any "\"" escape characters
note_vals <- coral_df %>% 
  select(starts_with("note")) %>% 
  gather() %>% 
  distinct() %>% 
  pull() %>%
  tolower() %>%
  str_replace_all("\"", "")

# Some of these values are actually multiple values separated by commas
# extract these values and consider them separately
note_vals <- note_vals %>% 
  str_split(",") %>% 
  unlist() %>% 
  str_trim() %>% 
  unique()

# Get all the values thath start with "fu"
# consolidate anything that is the pattern "fu1, fu2, fu3, ..." into just "fuX"
note_vals %>% 
  str_subset("^fu") %>%
  str_replace_all("fu[0-9]+", "fuX") %>% 
  unique()

# Get all values that start with "fi"
note_vals %>% 
  str_subset("^fi") %>% 
  str_replace_all("fi[0-9]+", "fiX") %>% 
  unique()

```


---





